Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-08-30 paper/ms.tex
updated methodology

Our method is based on the comparison of observations and mock
catalogs. We use two different kinds of statistics to perform the
comparison: (i) the distribution of the surface number density
across fields fields and (ii) the angular correlation function measured in
some fields.

@@

\subsection{Observational Constraints}

The first observational constraint we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arc-min, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter used in the survey is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a
total  $2161$  LAEs with an observed equivalent width, in the observer
frame, larger than $190$\AA 
over a total survey area of $2.42$ deg$^{2}$ that includes 12 sub-fields, 
this corresponds to average surface number density of $0.20\pm 0.01$
arcmin$^{-2}$.

@@

The second constraint is the angular correlation function
(ACF). \cite{Yamada2012} does not report an ACF measurement. Instead
we use the results by \cite{Hayashino2004} and
\cite{Ouchi2008,Ouchi2010}. \cite{Hayashino2004} observed in the
densest field of SSA22 while \cite{Ouchi2008} observed on a 1 deg$^2$
sky of the SXDS Field.  There are some differences between these
observations and those by \cite{Yamada2012}. The details in the color
selection, corresponding limiting luminosities and EW thresholds are
different in these references. Nevertheless we use these observations
as additional constraints in spite of the fact that the first selected
models are based only on the surface density statistics by
\citep{Yamada2012}.

\subsection{Simulation and Halo Catalogs}

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The catalogs were obtained from the publicly available
Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}}
\citep{MultiDark}. For each halo in the box we store its
position in the box (3-D coordinates) and FOF mass. We focus our work
on halos more massive than $1\times 10^{10}$\hMsun that are resolved
with at least $70$ particles, the reasons for this choice are
explained in the next sub-section.

\subsection{A Model to Populate Halos with LAEs}
\label{subsec:mocks}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\linewidth,angle=0]{./plots/Figure0.pdf}
\caption{ \label{fig:distros} Spatial distribution of a LAEs mock
  survey for a model with parameters $\log_{10}M_{\rm min}=10.4$, $\log_{10}M_{\rm
    max}=10.5$ and $f_{\rm occ}=0.1$ for the {\texttt match}
  method. The larger panel shows $7$ mock
  fields together mimicking the SSA22 region. The intermediate panel
  shows $3$ mock fields corresponding to the SXDS region. The lower
  panels represent the SDF and GOODS-North fields. The same model has
  another 14 associated mock surveys build from different regions in
  the simulation, but constructed following the same pattern in the
  {\texttt match} method.}
\end{center} 
\end{figure*}

We assume that a dark matter halo can only host one
detectable LAE at most.  There are three parameters that
decide whether a halo host a LAE: the lower and upper bounds for the
mass range $M_{\rm min}< M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm occ}$ of such halos that host a detectable LAE. The reader
must keep in mind that the physical interpretation of the occupation
fraction $f_{\rm occ}$ convolves two phenomena: the presence of a star
forming galaxy in a halo and its detectability as a LAE.  We do not
peform a explicit modeling for LAE detectability. Our model does not
assign a luminosity or escape fraction to each LAE.  We are only
interested in constraining the halo mass range hosting detectable LAEs
under the conditions defined by \cite{Yamada2012}.

@@

Next we construct mock surveys by making groups of $12$ mock fields
out of the $210$ available volumes. In total $15$ mock surveys are
constructed for each model $\mathcal{M}$. The grouping of the $12$
mock fields into a mock catalog is done in two different ways. The
first is called {\texttt match}, it follows the clustering of the
observational fields. From the $12$ mock fields, $7$ are constructed
from contiguous fields in the simulation to mimic the SSA22 region,
$3$ are also contiguous between them but not to the first $7$ fields
to mimic the SXDS fields and finally $2$ non-contiguous fields to
imitate the SDF and GOODS-North field.  The second way to group the
mock fields is called {\texttt   random}, whereby all the $12$ fields
are selected in such a way as to avoid that any two volumes are
contiguous. In this \documentname we only report the results obtained
by the {\texttt match} method and mention explicitly differences
observed with the {\texttt random} selection.

We note that we use $15\times 12=144$ mock fields out of the total of $210$
available sub-volumes. The reason is that the {\texttt match}
method imposes constraints on the way the $7$ fields mimicking the
SSA22 can be distributed. This restriction makes unusuable some of the
sub-volumes in the box. We decide to keep the number of mock surveys
fixed to $15$ also for the {\texttt random} method in order to allow a
fair comparison between the two methods.

\subsection{Exploring and Selecting Good Models}

We make a thorough exploration of the parameter space for the models
${\mathcal M}$. $\log_{10} M_{\rm min}$ takes $30$ values from $10.0$ up
to $12.9$ with an even spacing of $0.1$ dex. $\log_{10} M_{\rm max}$
takes values in the same range as $\log_{10}M_{\rm min}$ only with a
displacement of $0.1$ dex in the whole range. The occupation fraction
$f_{\rm occ}$ takes $10$ different values from $0.1$ to $1$ regularly
spaced by $0.1$. In total the number of different models ${\mathcal
  M}$ that are explored is $30 \times 30 \times 10 = 9000$.

The lower limit for the parameter $M_{\rm min}$ is set by the minium
occupation fraction we are able to consider. At $M_{\rm
  min}=10^{10}$\hMsun the halo number density is already $\sim 10$
times higher than the observational constraints for LAEs. This means
that models in that mass range and an occupation fraction $f_{\rm
  occ}=0.1$ have the possibility to be compatible with observations. Lower
values for $M_{\rm min}$ require $f_{\rm occ}<0.1$, which are not
considered in this \documentname.

@@

The second constraint  comes from the angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed  ( using 
the Landy \&  Szalay  estimator \citep{Landy1993} )  on two kinds of
fields in the mock survey: the densest and one corresponding to the
average density. The first with the \cite{Hayashino2004}  and \cite{Ouchi2010}
ACFs respectively.

The observational ACF and the  mock  ACF of the models are compared   
in terms of the correlation length ($\theta_{0}$) by fitting 
them to a power-law function:

\begin{equation}
\omega(\theta) = \left(\frac{\theta}{\theta_{0}}\right)^{-\beta}
\label{eq:fitting}
\end{equation}

Both mock and observational ACF are derived by a least square
minimization procedure. For each mock we obtain a covariance matrix
that gives us the uncertainty in the parameters $\beta$ and
$\theta_0$. We compute a mean value for the parameters and its
associated uncertainty as the cuadratic average of the uncertainties in
the fitting procedure. We consider that a model is consistent with
observations if the two parameters $\beta$ and $\theta_0$ are equal
within a $1$-$\sigma$ range.

>>>>

Our method is based on the comparison of observations and mock
catalogs. We use two different kinds of statistics to perform the
comparison: (i) the distribution of the surface number density
across fields and (ii) the angular correlation function measured in
some fields.

@@

\subsection{Observational constraints}

The first observational constraint we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arc-min, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter used in the survey is centered at $4977$ \AA~with  $77$ \AA~width,
corresponding to the redshift range $z=3.062$-$3.125$ and $41$ \hMpc comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a total  $2161$  LAEs with
an observed equivalent width, in the observer frame, larger than $190$
\AA~over a total survey area of $2.42$ deg$^{2}$ that includes 12
sub-fields,  this corresponds to average surface number density of
$0.20\pm 0.01$ arcmin$^{-2}$.

@@

The second constraint is the angular correlation function
(ACF). \cite{Yamada2012} does not report an ACF measurement. Instead
we use the results by \cite{Hayashino2004} and
\cite{Ouchi2008,Ouchi2010}. \cite{Hayashino2004} observed in the
densest field of SSA22 while \cite{Ouchi2008} observed on a $\sim 0.5$
deg$^2$ field.  There are some differences between these
observations and those by \cite{Yamada2012}. The details in the color
selection, corresponding limiting luminosities and EW thresholds are
different in these references. Nevertheless we use the ACF from these
observations as additional constraints in spite of the fact that the
first selected models are based only on the surface density statistics by
\cite{Yamada2012}.

\subsection{Simulation and halo catalogs}

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The catalogs were obtained from the publicly available
Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}}
\citep{MultiDark}. For each halo in the box we store its
comoving position in the box (3-D coordinates) and FOF mass. We focus our work
on halos more massive than $1\times 10^{10}$\hMsun that are resolved
with at least $70$ particles, the reasons for this choice are
explained in the next sub-section.

\subsection{A model to populate halos with LAEs}
\label{subsec:mocks}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\linewidth,angle=0]{./plots/Figure0.pdf}
\caption{ \label{fig:distros} Spatial distribution of a LAEs mock
  survey for a model with parameters $\log_{10}M_{\rm min}=10.4$, $\log_{10}M_{\rm
    max}=10.5$ and $f_{\rm occ}=0.1$ for the {\texttt{match}}
  method. The larger panel shows $7$ mock
  fields together mimicking the SSA22 region. The intermediate panel
  shows $3$ mock fields corresponding to the SXDS region. The lower
  panels represent the SDF and GOODS-North fields. The same model has
  another 14 associated mock surveys build from different regions in
  the simulation, but constructed following the same pattern in the
  {\texttt{match}} method.}
\end{center} 
\end{figure*}

We assume that a dark matter halo can only host one
detectable LAE at most.  There are three parameters that
decide whether a halo host a LAE: the lower and upper bounds for the
mass range, $M_{\rm min}< M_{\rm h} < M_{\rm max}$, where LAEs reside and the fraction $f_{\rm occ}$ of such halos that host a detectable LAE. The reader
must keep in mind that the physical interpretation of the occupation
fraction $f_{\rm occ}$ convolves two phenomena: the actual presence of a star
forming galaxy in a halo and its detectability as a LAE.  We do not
peform a explicit modeling for LAE detectability. Our model does not
assign a luminosity or escape fraction to each LAE.  We are only
interested in constraining the halo mass range hosting detectable LAEs
under the conditions defined by \cite{Yamada2012}.

@@

Next we construct mock surveys by making groups of $12$ mock fields
out of the $210$ available volumes. In total $15$ mock surveys are
constructed for each model $\mathcal{M}$. The grouping of the $12$
mock fields into a mock catalog is done in two different ways. The
first is called {\texttt{match}}, it follows the clustering of the
observational fields. From the $12$ mock fields, $7$ are constructed
from contiguous fields in the simulation to mimic the SSA22 region,
$3$ are also contiguous between them but not to the first $7$ fields
to mimic the SXDS fields and finally $2$ non-contiguous fields to
imitate the SDF and GOODS-North field.   The second way to group the
mock fields is called {\texttt{random}}, whereby all the $12$ fields
are selected in such a way as to avoid that any two volumes are
contiguous. In this \documentname we only report the results obtained
by the {\texttt{match}} method and mention explicitly differences
observed with the {\texttt{random}} selection.

Figure \ref{fig:distros} shows the spatial distribution for one mock
survey constructed using the {\texttt{match}} method. Each field
corresponds to one of the observational fields. The model parameters
to build it are $M_{\rm min}=10^{10.4}$\hMsun, $M_{\rm
  max}=10^{0.5}$\hMsun and $f_{\rm occ}=0.1$. It is only one out of
the $15$ different mock surveys that are constructed for each
model. We note that we use $15\times 12=180$ mock fields out of the
total of $210$ available sub-volumes. The reason is that the {\texttt{match}}
method imposes constraints on the way the $7$ fields mimicking the
SSA22 can be distributed. This restriction makes unusuable some of the
sub-volumes in the box. We decide to keep the number of mock surveys
fixed to $15$ also for the {\texttt{random}} method in order to allow a
fair comparison between the two methods.

\subsection{Exploring and selecting good models}

We make a thorough exploration of the parameter space for the models
${\mathcal M}$ where $\log_{10} M_{\rm min}$ takes $30$ values from $10.0$ up
to $12.9$ with an even spacing of $0.1$ dex. $\log_{10} M_{\rm max}$
takes values in the same range as $\log_{10}M_{\rm min}$ only with a
displacement of $0.1$ dex in the whole range. The occupation fraction
$f_{\rm occ}$ takes $10$ different values from $0.1$ to $1$ regularly
spaced by $0.1$. In total the number of different models ${\mathcal
  M}$ that are explored is $30 \times 30 \times 10 = 9000$.

The lower limit for the parameter $M_{\rm min}$ is set by the minimum
occupation fraction we are able to consider. At $M_{\rm
  min}=10^{10}$\hMsun the halo number density is already $\sim 10$
times higher than the observational constraints for LAEs. This means
that models in that mass range and an occupation fraction $f_{\rm
  occ}=0.1$ have the possibility to be compatible with observations. Lower
values for $M_{\rm min}$ require $f_{\rm occ}<0.1$, which are not
considered in this \documentname.

@@

The second constraint  comes from the angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed using 
the Landy \&  Szalay estimator \citep{Landy1993}  on two different 
fields in the mock survey. The first field is the densest to be compared
against the results of \cite{Hayashino2004}. The second field
has an average number density and is compared against \cite{Ouchi2010}.

The observed and mock ACF are fit to a power-law function:
\begin{equation}
\omega(\theta) = \left(\frac{\theta}{\theta_{0}}\right)^{-\beta}, 
\label{eq:fitting}
\end{equation}
where $\theta_0$ and $\beta$ are free parametres. The fit is donde
using a least square minimization procedure. For each mock field we obtain a
covariance matrix that gives us the uncertainty in the parameters $\beta$ and
$\theta_0$. Using the $15$ mock fields in a mock survey, we compute a
mean value for the parameters and its associated uncertainty as the
cuadratic average of the uncertainties in the fitting procedure. We
consider that a model is consistent with observations if the two
parameters $\beta$ and $\theta_0$ are equal within a $1$-$\sigma$
range.
